---
title: "1.wrangling-yield"
output: html_document
---
# Learning objectives  
Today's learning objectives are:  
- Talk about the role of yield monitor data in ag  
- Import spatial data  
- Learn and implement data wrangling steps  
- Learn and implement yield monitor data cleaning steps  
- Compare numerically and graphically raw and cleaned yield monitor data  


# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
```

```{r reading data}
yield17 

yield17
```

# Exploratory data analysis
```{r class}
class(yield17)
```

```{r summary}

```

What columns do you think are relevant?  

```{r yield distribution - density plot}


```

```{r speed distribution - density plot}


```

```{r yield map}


```

```{r yield restricted map}
ggplot(filter(yield17, Dry_Yield<5000) , aes(x=POINT_X,
                                             y=POINT_Y))+
  geom_point(aes(color=Dry_Yield))+
  scale_color_gradient(low="red", high = "green")

```

```{r yield x speed}


```

Do we need to clean this data?  
From a yield map quality perspective, what steps would you do?  
-   
-  

```{r data needs}
yield17
```

From a data wrangling perspective, what steps would you do?  
- standardize column naming style (for ease of use)  
- select only needed columns:  
- Speed  
- Dry_Yield  
- POINT_X  
- POINT_Y  

- rename columns to include measurement units  
- create columns in SI system  
- transform it into spatial object (as of now, it is only a data frame)  
- transform from unprojected to projected  

# Wrangling  
```{r yield17_w1 - clean_names}
# Standardizing names
yield17_w1 


yield17_w1
```

```{r yield17_w2 - select}
# Selecting needed columns
yield17_w2 


yield17_w2
```


```{r yield17_w3 - rename}
# Rename to include units
yield17_w3 

yield17_w3

```

```{r yield17_w4 - mutate}
# Creating yield and speed in SI
yield17_w4

# converting speed from mph to kmh: *1.09728
# converting yield from lb/ac to kg/ha: *0.453592/0.404686
yield17_w4
```

```{r yield17_w5 - st_as_sf}
# Creating spatial object
yield17_w5 

yield17_w5

class(yield17_w5)
```
CRS: NA? we need to fix that.  


```{r yield17_w6 - st_set_crs}
# Assigning native coordinate reference system EPSG identifier
yield17_w6 

yield17_w6
```
Geodetic CRS is fixed now. But what if we want it on a projected crs?  

```{r yield17_w7 - st_transform}
# Transforming from unprojected to projected (UTM NAD83 16 n)
yield17_w7

yield17_w7
```

Projected CRS fixed!  

```{r yield17_w - piping it all}
yield17_w <- yield17 %>% 
  clean_names() %>%
  dplyr::select(speed,
                dry_yield,
                point_x,
                point_y) %>%
  rename(speed_mph=speed,
         yield_lbac=dry_yield) %>%
  mutate(yield_kgha=yield_lbac*0.453592/0.404686,
         speed_kmh=speed_mph*1.09728) %>%
  st_as_sf(coords = c("point_x", "point_y")) %>%
  st_set_crs(value = 4326) %>%
  st_transform(crs = 6345)

yield17_w
```

# Yield Editor  
```{r yield map raw}
ggplot(yield17_w)+
  geom_sf(aes(color=yield_lbac), size=.1)+
  scale_color_gradient(low="red", high = "green")


```

Different shape than original plot, now it is a spatial object.  

Today we are applying 5 filters:  
  - Min and max yield  
  - Min and max speed  
  - Position  

## Min/max yield
```{r min max yield - density}
ggplot(data=yield17_w, aes(x=yield_lbac))+
  geom_density()

```

```{r yield summary}


```


```{r yield quantiles}
lowq_yield 
lowq_yield

upq_yield 
upq_yield

iqr <- upq_yield - lowq_yield
iqr

miny 
miny

maxy 
maxy
```

```{r min max yield annotated- density}
ggplot()+
  geom_density(data=yield17_w, aes(x=yield_lbac))

  #annotate(geom="rect",
  #         ymin=0, ymax=Inf,
  #         xmin=miny, xmax=maxy,
  #         alpha=.6, 
  #         fill="green")

```


```{r yield filtering}
yield17_w_yf 

yield17_w_yf
```
Before filtering we had a total of 31,724 data rows.  
How many rows now after filtering?  


```{r yield filtered - stats}
summary(yield17_w_yf$yield_lbac) 
```

All right, all done filtering based on yield min and max.  
Now let's do the same for speed.  

## Min/max speed  
```{r min max speed - density}
ggplot(data=yield17_w, aes(x=speed_mph))+
  geom_density()

```

```{r speed summary}
summary(yield17_w$speed_mph)
```


```{r speed quantiles}
lowq_speed <- quantile(yield17_w$speed_mph, 0.08)
lowq_speed

upq_speed <- quantile(yield17_w$speed_mph, 0.99)
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (0.18*iqr)
minv

maxv <- upq_speed + (0.18*iqr)
maxv
```


```{r min max speed annotated- density}
ggplot()+
  geom_density(data=yield17_w, aes(x=speed_mph))+
  geom_vline(xintercept = lowq_speed, color="red", linetype=2)+
  geom_vline(xintercept = upq_speed, color="blue", linetype=2)+
  geom_vline(xintercept = minv, color="red", linetype=1)+
  geom_vline(xintercept = maxv, color="blue", linetype=1)+
  annotate(geom="rect",
           ymin=0, ymax=Inf,
           xmin=minv, xmax=maxv,
           alpha=.6, 
           fill="green")

```

```{r speed filtering}
yield17_w_sf 

yield17_w_sf
```
Before filtering we had a total of 30,385 data rows.  
How many rows now after filtering?  


```{r speed filtered - stats}
summary(yield17_w_sf$speed_mph) 
```

# Position  
```{r importing boundary}
boundary 

boundary
```

```{r boundary_w}
boundary_w 
  
boundary_w
```

```{r boundary_w plot}

```

Let's create a negative buffer of 20-m.  
```{r buffer}
buffer 

buffer
```

```{r buffer plot}
yieldmap_raw+ 
  geom_sf(data=boundary_w, fill=NA)

```

```{r position filtering}
yield17_w_pf 

yield17_w_pf
```
Before filtering we had a total of 28,950 data rows.  
How many rows now after filtering?  

```{r position summary}
summary(yield17_w_pf$yield_lbac)
```

# Raw vs Cleaned data  
## Summary statistics
```{r raw summary}
yield17_w %>%
  summarise(
            )
```

```{r raw filtered}
yield17_w_pf %>%
  summarise(n=length(yield_lbac),
            min=min(yield_lbac),
            mean=mean(yield_lbac),
            max=max(yield_lbac),
            sd=sd(yield_lbac),
            cv=(sd/mean)*100
            )
```

## Density plot  
```{r}

```

# Map  
```{r raw map}
yieldmap_raw
```

```{r cleaned map}
yieldmap_clean <- ggplot(yield17_w_pf)+
  geom_sf(aes(color=yield_lbac), size=.1)+
  scale_color_gradient(low="red", high = "green")

yieldmap_clean
```

```{r side-by-side maps}
#install.packages("patchwork")
library(patchwork)


# saving to file

```

# Summary  
Today we did a lot! Here's a snapshot:  
- Imported a raw yield monitor data, explored it numerically and graphically  
- Wrangled this data set by implementing multiple steps  
- Learned about using the `%>%` to combine multiple steps into one overall workflow  
- Learned about and implemented 5 yield filters (yield, speed, position)  
- Compared raw and cleaned yield monitor data  



