---
title: "03-16_3.yieldstability"
output: html_document
---

# Learning Objectives  
Today, our learning objectives are to:  
- Import **interpolated cleaned yield data** for 3 years  
- Create year-specific spatial yield stability areas  
- Create spatio-temporal yield stability areas  
- Export the final yield stability layer  

# Setup  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(viridis)
library(ggthemes)
library(tibble)
#install.packages("tibble")

```

```{r yieldic17}
yieldic17

yieldic17

summary(yieldic17)
```


```{r yieldic19}
yieldic19 

yieldic19

summary(yieldic19)
```


```{r yieldic20}
yieldic20 

yieldic20

summary(yieldic20)
```
> What year was the highest yielding?  
> What about the lowest yielding?  

> Same number of observations for all layers?  
> Why is that?  

# Spatial stability  
Every time we deal with yield data across various years/fields/crops, there are challenges:  
- If different years/fields/crops, how to compare?  
- Even if same field but different years/crops, how to compare?  

Some solutions:  
- Standardize by mean/median/max (numerical). Ex. field mean in a given year was 1500 lb/ac, so divide all other yield values from that year by 1500.  

- Standardize by quantile (categorical). Ex. field median (e.g., 50% quantile) was 2000 lb/ac, so all observations below 2000 are classified as "low", and all above 2000 are classified as "high" for that year.  

**In our case, we are using the quantile approach.**
First let's calculate median yield per year, and then classify all yield observations as lower than the median (low) or higher than the median (high).  

Beginning with 2017 yield data.
```{r s_17}
s_17 

s_17
```


Let's check how does this classification look like in space.  
```{r s_17 map}
s_17_map 


s_17_map 
```

Repeat the same steps for 2019.  
```{r s_19}
s_19 <- yieldic19 %>%
  rownames_to_column(var = "id") %>%
  mutate(medianyield_lbac=median(ipyield_lbac)) %>%
  mutate(snyield=case_when(
    ipyield_lbac >= medianyield_lbac ~ "high",
    ipyield_lbac < medianyield_lbac ~ "low"
  )) %>%
  dplyr::select(id, snyield_19=snyield, geometry) 

s_19
```

```{r s_19 map}
s_19_map <- ggplot()+
  geom_sf(data = s_19,
          aes(fill=snyield_19), color=NA)+
  scale_fill_viridis_d(option="E")+
    labs(title="2019")+
  theme_map()+
  theme(legend.position = "none")


s_19_map 
```
Repeat the same steps for 2020.  
```{r s_20}
s_20 <- yieldic20 %>%
  rownames_to_column(var = "id") %>%
  mutate(medianyield_lbac=median(ipyield_lbac)) %>%
  mutate(snyield=case_when(
    ipyield_lbac >= medianyield_lbac ~ "high",
    ipyield_lbac < medianyield_lbac ~ "low"
  )) %>%
  dplyr::select(id, snyield_20=snyield, geometry) 

s_20
```

```{r s_20 map}
s_20_map <- ggplot()+
  geom_sf(data = s_20,
          aes(fill=snyield_20), color=NA)+
  scale_fill_viridis_d(option="E")+
  labs(title="2020")+
  theme_map()+
  theme(legend.position = "none")

s_20_map 
```

Let's plot all 3 maps side-by-side to assist in visualizing them.  
```{r all maps}
library(patchwork)


```



# Spatio-temporal stability  
To perform a temporal analysis, we need to merge all three data frames into one.  

After joining the data sets, we still only have their information on year-specific spatial classification (low and high yielding). 

> How can we merge these pieces of information to create a spatial-temporal classification?  

Here's how I want to classify our spatial-temporal classes:  
- If a pixel was low yielding in all 3 years, then it gets classified as stable low (sl)  

- If a pixel was high yielding in all 3 years, then it gets classified as stable high (sh)  

- Otherwise, then it gest classified as unstable (uns).  


```{r st_all}
st_all 

st_all
```


```{r st_all_map}
ggplot()+
  geom_sf(data = st_all,
          aes(fill=stnyield), color=NA)+
    scale_fill_viridis_d(option="C")+
  theme_map()

```
> Just eyeballing, what proportion of this field do you think is unstable? 20%? 50%?  






Let's calculate below:  
```{r proportion per stability zone}


```
Imagine this is your farm, and I tell you that in about 43% of your field, yield can be among highest or lowest just depending on the year.  

> How would you manage it?  


# Exporting  
Let's export this data set so we can use it again in a future exercise.  
```{r exporting yield stability}


```

# Summary  

Today we:  
- Imported all three interpolated cleaned yield data  
- Classified each year-specific yield pixel as low vs. high (spatial stability)  
- Joined the three data sets, and classified each pixel as stable high, stable low, or unstable (spatial-temporal stability)  
- Exported the spatial-temporal stability layer   

