---
title: "04-25_7.VRN_completed"
output: html_document
---
# Learning objectives  
The learning objectives for this exercise are to:  
  - Learn where to find crop-specific nutrient recommendations  
  - Understand what information is needed to recommend N in cotton   
  - Estimate zone-specific cotton yield potential  
  - Calculate zone-specific total N recommendation 
  - Calculate zone-specific side-dress variable N rate recommendation  
  - Calculate total amount of fertilizer to order  
  - Export recommendation to file to be used in a variable rate controller  

# Introduction  
We are going to create a recommendation to apply variable rate nitrogen (N) fertilizer in cotton, with a different in-season side-dress rate for each zone.    

> How do we know how much N cotton needs (or really, how do we know how much of any given nutrient any crop needs)?  

A few sources of this type of information would be **research papers** and **extension publications**.  

Normally (in the US, at least), crop-specific nutrient recommendations are available through extension publications.

A quick Google search on **"cotton nutrient recommendation in Georgia"** (notice the specificity of crop and state in the key words, important as each state would have their own recommendation algorithm for a given crop) returns a few results.

For Georgia, UGA publishes annually the **"Cotton Production Guide"**, which contains not only nutrient but all sorts of management recommendations for cotton. The UGA 2022 Cotton production guide can be found [here](https://secure.caes.uga.edu/extension/publications/files/pdf/AP%20124-2_1.PDF).

Pages 9 and 10 of the guide have information on N in cotton, including a table relating yield potential with N rate.  

This is the same information as what we saw with Dr. Virk on the tool [UGAFertEx](https://aesl.ces.uga.edu/calculators/ugfertex/).  


From that table and accompanying text on the guide, we can see that cotton total N rate:  
- Depends on **yield potential**  
- Should be **adjusted in +25%** in certain cases (e.g., cotton following cotton)    
- Should be **split-applied**  

> So, what are the required pieces of information that we need to know in order to make a recommendation of N in cotton?  

1.  
2.  
3.  

> How can we define yield potential for each of the zones?  

# Setup  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(dplyr)
library(tidyr)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "right",
        panel.background = element_rect(fill="gray80",
                                        color=NA),
        plot.background = element_rect(fill="gray80",
                                       color=NA),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

```


```{r yieldic17}
yieldic17 <- read_sf("../data/KB2_Yield_2017_clean_ip.geojson") %>%
  mutate(yield17_lbac=ipyield_lbac*.4) %>%
  dplyr::select(yield17_lbac, geometry)

yieldic17

summary(yieldic17)
```


```{r yieldic19}
yieldic19 <- read_sf("../data/KB2_Yield_2019_clean_ip.geojson") %>%
  dplyr::select(yield19_lbac=ipyield_lbac, geometry)

yieldic19

summary(yieldic19)
```
```{r yieldic20}
yieldic20 <- read_sf("../data/KB2_Yield_2020_clean_ip.geojson") %>%
  dplyr::select(yield20_lbac=ipyield_lbac, geometry)


yieldic20

summary(yieldic20)
```

```{r zone_st}
zone_st <- read_sf("../data/zone_st.geojson") %>%
  dplyr::select(zone, geometry)

zone_st
```

# Wrangling  
```{r Merging yield and zones}
zone_yield <- zone_st %>%
  st_join(yieldic17,
          join = st_equals, 
          left = T
  ) %>%
  st_join(yieldic19,
          join = st_equals, 
          left = T
  ) %>%
  st_join(yieldic20,
          join = st_equals, 
          left = T
  ) 

zone_yield
```
# VRN  
Below, we will
- Calculate zone-specific yield potential (as the mean of 3 years)    
- Calculate the total N rate (lbs N/ac) needed with the algorithm for N in cotton from the production guide
- Adjust it in +25% assuming that we are following cotton  
- Adjust it in 2/3 to be applied side-dress (1/3 applied at planting)  
- Define our N source as urea-ammonium nitrate (UAN) 28% N  
- Calculate the total product rate (gal UAN/ac)  


```{r vrn calculation }
vrn <- zone_yield %>%
  pivot_longer(-c(zone, geometry)) %>%
  group_by(zone) %>%
  # Defining yield potential  
  summarise(yieldpotential=quantile(value,.9)) %>%
  mutate(totalnrate_lbsac=15 + yieldpotential*0.06) %>% 
  # Adjust total rate in 25% upwards for following cotton
  mutate(totalnrate_adj_lbsac=totalnrate_lbsac*1.25) %>%
  # Calculate pre-plant rate
  mutate(pp_totalnrate_adj_lbsac=totalnrate_adj_lbsac*.33) %>%
  # Calculate ~2/3 as portion to be applied in-season (Assuming a single pre-plant rate was applied to both zones)
  mutate(is_totalnrate_adj_lbsac=totalnrate_adj_lbsac-48) %>%
  # Calculate product rate when using UAN solution 28% N
  # UAN 28% has 3 lbs N per gallon of product (see chunk below)
  mutate(uanrate_galac=is_totalnrate_adj_lbsac/3,
         uanrate_galac=round(uanrate_galac,1))

vrn
```
UAN is a liquid solution, so how do we go from volume of solution to lbs of N?  

> What information do we need?  










We can find this information on the fertilizer label/msds information from manufacturer.  

Normally, UAN 28% density is 10.67 lbs/gal, as we can find from these retailers:  
- https://www.agricocanada.com/wp-content/uploads/2016/09/agric-101-SDS_Urea-Ammonium-Nitrate-UAN-Solution.pdf  
-  https://www.cfindustries.com/globalassets/cf-industries/media/documents/safety-data-sheets/uan---north-america/urea_ammonium_nitrate_solution_uan_sds.pdf  

NOTE: UAN is normally sold as 28, 30, or 32%N. Each of them has a different density!  


```{r UAN volume to mass of N}
uan28density_lbsgal <- 10.67

uan28Nconc_pct <- 0.28

uan28_lbsNgal <- uan28density_lbsgal * uan28Nconc_pct

uan28_lbsNgal #~3 lbs N/gal of UAN28%

```

How much UAN28% do we need to have delivered at the field to ensure enough product for application?  

```{r total volume of UAN per area}
vrn %>%
  mutate(area_m2=st_area(.)) %>%
  mutate(area_ac=as.numeric(area_m2/4047)) %>%
  # Will ask for 20% more than recommended to have a bit extra 
  mutate(uan_gal=(uanrate_galac*area_ac)*1.2) %>%
  summarise(totaluan_gal=sum(uan_gal),
            totaluan_gal=round(totaluan_gal,0))

```

# Rx map  
```{r vrn vector}
vrn_v <- zone_yield %>%
  left_join(vrn %>%
              st_drop_geometry()
  )


vrn_v
```

```{r vrn plot}
vrn_v %>%
  mutate(uanrate_galac=factor(uanrate_galac)) %>%
  ggplot()+
  geom_sf(aes(fill=uanrate_galac,
              color=uanrate_galac))+
  labs(title="VRN Rx - UAN 28% rate",
       fill="UAN rate\n(gal/ac)",
       color="UAN rate\n(gal/ac)")+
  mytheme_map

ggsave("../output/VRN.png",
      width = 4, 
       height = 4
       )  

```

# Exporting  
We are going to export as **shapefile** as it is the most commonly accepted format in variable rate controllers.  

Since we are exporting as shapefile, let's create a sub-folder inside `output` called `VRN` to have all the associated files saved together.  

```{r exporting}
vrn_v %>%
  dplyr::select(zone, uanrate_galac) %>%
  write_sf("../output/VRN/vrn.shp",
           delete_dsn = T)
```

# Summary  
In summary, today we:  
- Understood where to look for information on crop and nutrient-specific recommendations  
- Defined yield potential for each zone  
- Calculated VRN in both lbs N/ac and gallons UAN38/ac  
- Exported VRN Rx as a shapefile  

Next, we will wrap up and perform a profitability analysis.  




